Distribution-oblivious Programming with Dish

### 2019-10-09 pldi/draft.tex
Update intro

Combining several features, \unix makes small tasks easy to express:
  $p_1$ creates a character stream, breaks it down to words, transliterates them lower-case, sorts them to group duplicates, counts and replaces duplicates, sorts them numerically and in reverse order, picks the top five results and writes them to a file \ttt{out}.
For a user with \emph{one} computer and small input, this pipeline takes a few seconds to compose and execute~\cite{bentley1986literate}.
% composition allows the composition of general primitives

@@

For a user with many computers and larger inputs, however, developing $p_1$'s distributed equivalent requires a significant effort.
% The scope of such rewrites, and therefore the cost of manual effort, can vary considerably.
In simple cases comprised of pure functions, has to be rewritten in the language and interfaces exposed by the chosen big-data framework or DSL.
This cost is unjustifiable for one-off pipelines that take a few minutes to compose (but are applied to large datasets).
More complex pipelines, such as the ones presented in later sections, require full-fledged distributed programming language or a distributed operating system---\eg Plan9's \ttt{rc} shell.
In these cases, manual rewriting is expensive and can introduce new bugs, cascading changes, or divergence from legacy functionality---
Could the generation and execution of $p_1$'s distributed version be fully and correctly automated?

@@

The key insight behind our work is that \unix pipelines already provide most of the information for distributing a computation.
Indeed, the primary reason behind $p_1$'s succinctness is that the pipeline is a domain-specific language for describing operations over streams.
Primitives 
\sys enables \emph{distribution-oblivious programming}: 
  multi-order of magnitude speedups with correctness guarantees and without any developer effort.

The a series of transformations that 
Second, we are using describing composition . For this, we are restricting the study to a subset of the POSIX shell--

Local operations and identifiers have to be carefully translated to distributed ones, a translation that depends on the types of the operations and associated identifiers.

The semantics is enough to capture powerful features such as stream branching and feedback constructs found in practice and in the system's evaluation.

The paper is structured as follows.
We first start with a background section outlining pipeline concepts.
Sections \ref{profiling}--\ref{recipes} highlight our key contributions:
\begin{itemize}

  \item
  \S\ref{overview} overviews the system and introduces the notion of command classes.
	introduces \emph{load-detection transformations} that collect windowed statistics about load at each module boundary.
  Control-plane coordination with a global view of load and available resources helps decide when to initiate scale-out of a bottlenecked module.

  \item
  \S\ref{core} presents a set of parametrizable \emph{distribution transformations} that transparently scale modules out.
  These transformations can create module replicas, hook communication channels among them, schedule requests, and forward side-effects such as mutation and collection of memory.

  \item
  \S\ref{recipes} outlines \emph{distribution recipes}, lightweight annotations that guide the semantics of the resulting distribution.
  They offer significant flexibility by\eat{ generating configurations that} parametrizing transformations, including tuning state management, replication consistency, event propagation, and colocation preferences.
\end{itemize}

\noindent
\sys's evaluation~\sx{eval} uses a combination of one-liner micro-benchmarks that highlight certain features and three multi-line programs---a web crawling and indexing engine, a production genomics pipeline, and a climate data processing.
We close with a discussion of related prior work~\sx{related} and possible future directions~\sx{discussion}.

>>>>

\noindent
Program $p_1$ creates a character stream, breaks it into words, transliterates to lower-case, sorts to group duplicates, summarizes them with a counter, sorts them in descending order, picks the top five results and writes them to a file \ttt{out}.
Combining several features, \unix makes small tasks easy to express;
  for a user with \emph{one} computer and small input, this pipeline takes a few seconds to compose and execute~\cite{bentley1986literate}.
% composition allows the composition of general primitives

@@

For a user with many computers and larger inputs, however, developing $p_1$'s distributed equivalent requires a significant effort.
% The scope of such rewrites, and therefore the cost of manual effort, can vary considerably.
For simple pipelines that capture in restricted models of computation, this effort amounts to expressing the computation using the primitives provided by a big-data framework~\cite{mapreduce:08, ciel:11, spark:12, naiad:13} or domain-specific language~\cite{alvaro2011consistency, distal:13, meiklejohn2015lasp}---an unjustifiable cost for one-off pipelines that take a few minutes to compose (but are applied to large datasets).
More complex pipelines, such as the ones presented in later sections, require full-fledged distributed programming language~\cite{erlang:96, lopes1997d, acute:05, mace:07, cloudhaskell:11, ScalaLoci:18}. %  or a distributed operating system---\eg Plan9's \ttt{rc} shell.
In these cases, manual rewriting is expensive and can introduce new bugs, cascading changes, or divergence from legacy functionality.
Could the generation and execution of $p_1$'s distributed version be fully and correctly automated?

@@

The key insight behind \sys is that the language of the Unix shell already encodes stream processing, providing most of the information required for distributing a computation.
\sys builds on this insight with a careful study of the scalability properties of shell primitives and its commands, paired with a novel rewriting scheme that identifies maximal pipeline sub-expressions that are candidates for scale-out.
We term the result \emph{distribution-oblivious programming}---a series of techniques that fully automates program distribution while provably maintaining correct (non-distributed) semantics.

Key benefits include the ability to scale legacy computations onto
\sys converts legacy shell pipelines into their distributed equivalents fully automatically, offering $100\times$ improvements in performance without a single line of additional code or annotation.
Most importantly, \sys provides an architectural lesson:
  the current design shows that large-scale efforts in the distributed- and operating-system literature to provide a \unix-like distributed equivalent could have been simplified by a thin (but sophisticated) rewriting shim like the one \sys provides.

% % Make Unix benefits explicit?
% Indeed, the primary reason behind $p_1$'s succinctness is that the pipeline is a domain-specific language for describing operations over streams.
% Key elements of \unix are the ability to compose programs written in different languages, the abstraction of a file system as a set of resident streams, a small but extensible library of commands, and the ability to resolve names within a global context.
% Under the hood, the \unix kernel buffers results, synchronizes processing stages, and generally orchestrates the computation.

% Primitives 
% \sys enables \emph{distribution-oblivious programming}: 
%   multi-order of magnitude speedups with correctness guarantees and without any developer effort.
% 
% The a series of transformations that 
% Second, we are using describing composition . For this, we are restricting the study to a subset of the POSIX shell--
% 
% Local operations and identifiers have to be carefully translated to distributed ones, a translation that depends on the types of the operations and associated identifiers.
% 
% The semantics is enough to capture powerful features such as stream branching and feedback constructs found in practice and in the system's evaluation.

The paper is structured as follows.
We first start with a background section outlining pipeline concepts~\sx{bg}.
Sections \ref{classes}--\ref{recipes} highlight our key contributions:\footnote{
  This is not precise, we'll need to wait until it matures a bit.
}
\begin{itemize}

  \item
  \S\ref{classes} overviews \sys and introduces several scalability and distributability classes.

  \item
  \S\ref{rewrite} presents a set of rewriting transformations for identifying maximal pipeline sub-expressions, candidates for scale-out.

  \item
  \S\ref{other} details other concerns, such as distributed state management and extensibility.
\end{itemize}

\noindent
\sys's evaluation~\sx{eval} uses a combination of one-liner micro-benchmarks that highlight certain features and three multi-line macro-benchmarks---a web crawling and indexing engine, a production genomics pipeline, and a large-scale weather analysis.
We close with a discussion of related prior work~\sx{related} and possible future directions~\sx{discussion}.
